---
title: "a4_task2_karla_garibay"
author: "Karla Garibay Garcia"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(effsize)
library(broom)
library(ggbeeswarm)
library(kableExtra)
```

### 1. Introduction

This report explores the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 in Bonanza Creek Experimental Forest. 


### 2. Data and analyses


### 3. Exploratory Findings

#### A. Annual juvenile hard trap counts


The total number of juvenile hare trappings during each year of the study, from 1999 - 2012, for all grid sites were counted. 


```{r}
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))

hares_counts <- bonanza_hares %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  filter(age == "j") %>% 
  count(year)

ggplot(data = hares_counts, 
       aes(x = year,
           y = n)) +
  geom_bar(stat = "identity", 
           fill = "cyan4") +
  labs(x = "Year", 
       y = "Count", 
       title = "Juvenile hare trappings from 1999-2012") +
  scale_x_continuous(breaks = seq(from = 1999, to = 2012, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 130, by = 10)) +
  theme_grey()
```


#### B. Visualize juvenile hare weights

#### C. Juvenile male & female snowshoe hare weight comparison 

#### D. Relationship between juvenile weight & hind foot length

### 4. Summary


### 5. Citations